---
title: "Table for Twitter during season"
author: "Harrison Stanton"
date: "2024-05-21"
output: html_document
---

```{r, message=FALSE}
library(nflfastR)
library(tidyverse)
library(dplyr)
library(nflreadr)
library(knitr)
library(kableExtra)
```

```{r, message=FALSE}
pbp_2023 <- load_pbp(2023)
roster_2023 <- fast_scraper_roster(2023)
```

```{r}
#create a new column that has who caught or rushed in the same column
pbp_2023 <- pbp_2023 %>%
  mutate(
    rush_receiver_id = case_when(
      is.na(receiver_player_id) & is.na(rusher_player_id) ~ 'NA',
      is.na(receiver_player_id) ~ rusher_player_id,
      TRUE ~ receiver_player_id
    )
  )
```

#back to what I most recently wrote
```{r}
roster_2023 <- roster_2023 %>% select(position, full_name, gsis_id, team)
roster_2023 <- roster_2023 %>% filter(position == "RB")
```

```{r}
pbp_2023 <- pbp_2023 %>%
  filter(rush_receiver_id %in% roster_2023$gsis_id)

pbp_2023 <- pbp_2023 %>%
  drop_na(rush_receiver_id)
```


```{r}
#this factors out targets that weren't caught 
pbp_2023 <- pbp_2023 %>%
  filter(!(is.na(receiving_yards) & !(is.na(receiver_player_id)))) %>%
  filter(is.na(two_point_conv_result)) %>%
  filter(season_type == "REG")
```


```{r}
#counting the games each RB played in 
games_df <- pbp_2023 %>%
  group_by(rush_receiver_id) %>%
  summarise(games = n_distinct(game_id))
```

```{r}
games_df <- games_df %>%
  mutate(gsis_id = rush_receiver_id)
```


```{r}
#this is where my problem is at... if a guy plays special teams or doesn't record a touch, he played a game but won't show up on my ish bc he didn't register a stat
games_maybe <- left_join(roster_2023, games_df, by = "gsis_id")
```


```{r}
tt_all <- pbp_2023 %>%
  group_by(receiver_player_id, rusher_player_id) %>%
  filter(is.na(two_point_conv_result)) %>%
  filter(!(is.na(receiving_yards) & !(is.na(receiver_player_id)))) %>%
  summarize(Carries_or_catches = n())


hvt_all <- pbp_2023 %>%
  group_by(receiver_player_id, rusher_player_id) %>%
  filter((!(is.na(rusher_player_id)) & yardline_100 <= 10) | !(is.na(receiver_player_id))) %>%
  summarize(High_value_touches = n()) 
```

```{r}
teams <- roster_2023 %>%
  distinct(team, .keep_all = TRUE) %>%
  pull(team)
```

```{r}
tt_all$receiver_player_id <- ifelse(is.na(tt_all$receiver_player_id), tt_all$rusher_player_id, tt_all$receiver_player_id)

tt_all$rusher_player_id <- ifelse(is.na(tt_all$rusher_player_id), tt_all$receiver_player_id, tt_all$rusher_player_id)

tt_all <- tt_all %>%
  group_by(receiver_player_id) %>%
  mutate(Total_touches = sum(Carries_or_catches)) %>%
  ungroup()
```

```{r}
tt_all <- tt_all %>%
  distinct(receiver_player_id, .keep_all = TRUE)


tt_all <- tt_all %>%
  select(receiver_player_id, Total_touches)

tt_all <- tt_all %>% rename(gsis_id = receiver_player_id)
```

```{r}
hvt_all$receiver_player_id <- ifelse(is.na(hvt_all$receiver_player_id), hvt_all$rusher_player_id, hvt_all$receiver_player_id)

hvt_all$rusher_player_id <- ifelse(is.na(hvt_all$rusher_player_id), hvt_all$receiver_player_id, hvt_all$rusher_player_id)

hvt_all <- hvt_all %>%
  group_by(receiver_player_id) %>%
  mutate(Total_HVT = sum(High_value_touches)) %>%
  ungroup()
```

```{r}
hvt_all <- hvt_all %>%
  distinct(receiver_player_id, .keep_all = TRUE)


hvt_all <- hvt_all %>%
  select(receiver_player_id, Total_HVT)

hvt_all <- hvt_all %>% rename(gsis_id = receiver_player_id)
```

```{r}
#High value + total touch data frame
hvt_tt_df <- left_join(roster_2023, tt_all, by = "gsis_id")

hvt_tt_df <- left_join(hvt_tt_df, hvt_all, by = "gsis_id")
#trying to find an efficient way to get number of games played in during the season into this thing... not sure how to do that
#I think I need to start by adding another column that's rusher_or_receiver,
#then do something like if(rusher_or_receiver_id) %in% rb_df (a list of all running backs), then keep it
#then group_by(rusher_or_receiver_id, week) and count the number of weeks that they played in... hoping this last step works
```

```{r}
hvt_tt_df <- left_join(hvt_tt_df, games_maybe %>% select("gsis_id", "games"), by = "gsis_id")
```

```{r}
hvt_tt_df <- hvt_tt_df %>% 
  mutate(
    HVT_per_game = Total_HVT / games,
    Touches_per_game = Total_touches / games,
    TRAP = (Total_touches - Total_HVT) / Total_touches
           )

hvt_tt_df <- hvt_tt_df %>% drop_na(Total_touches, Total_HVT)
```

```{r}
#random limit I decided... the problem that I run into sometimes is guys who play in more games, but don't get touches in all of them. For instance, Kyle Jus... he played in like 16 or 17 but only touched the ball in like 10, which makes his per game numbers a little off
#this filter is to try and make sure that all the people who have a game count on here actually played in that many games... I needed to use touches/game because I want to update this each week, so I can't just use how many games they play in or touches they have
#I think I could add to the filter what game of the season it is too... maybe like a max(week) column, then games played / total games >= .75 might help this too... not sure though tbh, this is a little jank 
hvt_tt_df <- hvt_tt_df %>% filter(Touches_per_game >= 8)
```

#stuff I pasted, trying to figure out what I need/don't
```{r}
pbp_2023_week1 <- pbp_2023 %>%
  filter(week == 1)
```

```{r, message=FALSE}
tt_week <- pbp_2023_week1 %>%
  group_by(receiver_player_id, rusher_player_id) %>%
  filter(is.na(two_point_conv_result)) %>%
  filter(!(is.na(receiving_yards) & !(is.na(receiver_player_id)))) %>%
  summarize(Carries_or_catches = n())


hvt_week <- pbp_2023_week1 %>%
  group_by(receiver_player_id, rusher_player_id) %>%
  filter((!(is.na(rusher_player_id)) & yardline_100 <= 10) | !(is.na(receiver_player_id))) %>%
  summarize(High_value_touches = n()) 
```


```{r}
tt_week$receiver_player_id <- ifelse(is.na(tt_week$receiver_player_id), tt_week$rusher_player_id, tt_week$receiver_player_id)

tt_week$rusher_player_id <- ifelse(is.na(tt_week$rusher_player_id), tt_week$receiver_player_id, tt_week$rusher_player_id)

tt_week <- tt_week %>%
  group_by(receiver_player_id) %>%
  mutate(Total_touches = sum(Carries_or_catches)) %>%
  ungroup()
```

```{r}
tt_week <- tt_week %>%
  distinct(receiver_player_id, .keep_all = TRUE)


tt_week <- tt_week %>%
  select(receiver_player_id, Total_touches)

tt_week <- tt_week %>% rename(gsis_id = receiver_player_id)
```

High value touches section:
```{r}
hvt_week$receiver_player_id <- ifelse(is.na(hvt_week$receiver_player_id), hvt_week$rusher_player_id, hvt_week$receiver_player_id)

hvt_week$rusher_player_id <- ifelse(is.na(hvt_week$rusher_player_id), hvt_week$receiver_player_id, hvt_week$rusher_player_id)

hvt_week <- hvt_week %>%
  group_by(receiver_player_id) %>%
  mutate(Total_HVT = sum(High_value_touches)) %>%
  ungroup()
```

```{r}
hvt_week <- hvt_week %>%
  distinct(receiver_player_id, .keep_all = TRUE)


hvt_week <- hvt_week %>%
  select(receiver_player_id, Total_HVT)

hvt_week <- hvt_week %>% rename(gsis_id = receiver_player_id)
```

```{r}
#High value + total touch data frame
hvt_tt_week <- left_join(roster_2023, tt_week, by = "gsis_id")

hvt_tt_week <- left_join(hvt_tt_week, hvt_week, by = "gsis_id")
```

```{r}
hvt_tt_week <- hvt_tt_week %>% filter(!is.na(Total_touches))
```

```{r}
hvt_tt_week$Total_HVT <- ifelse(is.na(hvt_tt_week$Total_HVT), 0, hvt_tt_week$Total_HVT)
```

```{r}
hvt_tt_week <- hvt_tt_week %>% rename(touches_week1 = Total_touches, HVT_week1 = Total_HVT)
```

#final table to output
```{r}
#the games is still a little messed up on this... gotta check that before I tweet this out 
#I don't think there's a perfect way to do this tbh... I might just need to put how many they have this week and how many total
output_df <- left_join(hvt_tt_df, hvt_tt_week %>% select(gsis_id, touches_week1, HVT_week1), by = "gsis_id")
output_df <- output_df %>% arrange(desc(HVT_week1))
output_df <- output_df %>% select(full_name, team, HVT_week1, touches_week1, Total_HVT, Total_touches, TRAP,  games)
```

```{r}
#just checking to see that my output_df is accurate
ekeler_pbp_w1 <- pbp_2023 %>%
  filter(rush_receiver_id == "00-0033699" & week == 1) %>%
  filter((play_type == "run" & yardline_100 <= 10) | play_type == "pass")

```

Another thing I want to figure out tomorrow is looking at Pass Rate Over Expectation (PROE). In column 371 and 372, they have xpass, which is the probability that a team passes on that play, and then pass_oe, which was -xpass if they ran the ball, and 1-xpass if they passed. I think then their PROE is just the sum of the pass_oe. I have to look this up though to be sure, and compare with people who already calculated these numbers. 

```{r}
pbp_2023_week1 %>% filter(!(play_type == "run" | play_type == "pass"))
```


```{r}
#this is just wrong... not quite right, and I'm not sure why tbh.
#gotta figure this out in the morning... this would be a dope piece to add to my tweets
#I think I might've missed something in the filter statements... no way that every team was below expectation bruh
#maybe I didn't filter out kickoffs? or extra points? or something like that jacked this all up
total_pass_oe <- pbp_2023 %>%
  filter(posteam == "BAL") %>%
  summarize(total_pass_oe = mean(pass_oe, na.rm = TRUE)) %>%
  pull(total_pass_oe)

# Print the result
print(total_pass_oe)
```

